---
title: "Internet Access in Illinois"
author: "Danya Sherbini"
date: '2022-08-30'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidycensus)
library(viridis)
```
# Introduction

## Using Tidycensus
I'm going to examine internet access data from the U.S. Census. Below I activate the tidycensus API by using my own census key. Once activated, I identified some variables of interest related to household internet access, such as demographic and income data.

```{r, message = FALSE, warning = FALSE}

census_api_key("338b89135dc1066f5241fa4e16804fc7345d1822", 
               install = TRUE, 
               overwrite = TRUE)
readRenviron("~/.Renviron")

v1 <- load_variables(2020, "acs5") #here I loaded variables for the 2020 ACS
```

## Loading Internet Access Data
I'm loading data from the 2020 American Community Survey and filtering by Illinois counties.
```{r, message = FALSE, warning = FALSE}

internet_data_raw <- get_acs( #calling data from the 5-year ACS
  geography = "county", #geo by county
  variables = c(with_subscription = "B28002_002", #HHs with internet sub
                without_subscription = "B28002_012", #HHs w internet but no sub
                no_internet = "B28002_013", #HHs w/o internet
                hhs_total = "B28002_001", #total households
                pop_total = "B01003_001", #total population
                num_white = "B02001_002", #number white
                num_black = "B02001_003", #number Black
                num_asian = "B02001_005", #number Asian
                num_latinx = "B03003_003",#number Latinx
                median_hhs_income = "B19013_001"), #Median household income
  year = 2020, #filtering by year 2020
  state = "IL" #filtering by IL
)

```

## Examining the Dataframe

```{r, message = FALSE, warning = FALSE}

nrow(internet_data_raw)
ncol(internet_data_raw)
names(internet_data_raw)

```

# Data Cleaning & Manipulation

```{r, message = FALSE, warning = FALSE}

internet_data <- internet_data_raw %>%
  select(-moe) %>% #removing moe column
  spread(variable, estimate) #transforming these variables from rows to columns


#adding new columns for %
internet_data <- internet_data %>%
  mutate(perc_nointernet = no_internet / hhs_total, 
         perc_withsubscription = with_subscription / hhs_total, 
         perc_wo_subscription = without_subscription / hhs_total, 
         perc_asian = num_asian / pop_total, 
         perc_black = num_black / pop_total, 
         perc_latinx = num_latinx / pop_total, 
         perc_white = num_white / pop_total)

```

# Data Visualization
The data viz below looks at the correlation between household median income, percent of households without internet access, and percent of Black residents by county.
```{r}

internet_data %>%
  ggplot(aes(x = perc_nointernet, y = median_hhs_income,
             color = perc_black)) +
  geom_point(size = 3) +
  scale_color_gradient(low="blue", high="red") +
  labs(title="Household Internet Access and Income in Illinois", x="Percent of Households w/o Internet Access", y="Median Household Income")+
  theme_classic()

```

